---
title: "04.2 - Poisson-distribution NE Wells"
author: "Robert C Cline Sr"
date: "April 18, 2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE) 
library(kableExtra)
library(tidyverse)
library(here)
library(readr)
library(tidyverse)
library(here)  
library(lubridate)
library(RODBC) 
library(janitor)
library(stringr) 
library(gt) 

```

```{r, include=FALSE} 
# source(here("scripts", "theme.clean.R", local = knitr::knit_global()))
# source("../scripts/theme.clean.R", local = knitr::knit_global())
# or sys.source("your-script.R", envir = knitr::knit_global())
``` 

```{r}  
# pathtallgrass <- "E:/tallgrass/data/NebraskaWellData_Paleozoic_Working.accdb"
# accdbfetch("E:/tallgrass/data/NebraskaWellData_Paleozoic_Working.accdb")

# accdbfetch::accdb_fetch(pathtallgrass) 

dbpath <- "E:/tallgrass/data/NebraskaWellData_Paleozoic_Working.accdb"

ch0<- RODBC::odbcDriverConnect(paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)};DBQ=",dbpath ))

welldata0 <- RODBC::sqlFetch(ch0, "tblNebraskaWellData")
tops <- RODBC::sqlFetch(ch0, "tblNebraskaWellData_Tops")
legend <- RODBC::sqlFetch(ch0, "tbl_Legend")
# welltype <- RODBC::sqlFetch(ch0, "Well_Type")
statuscode <- RODBC::sqlFetch(ch0, "Well_Status_Codes")
fmtn <- RODBC::sqlFetch(ch0, "tbl_Fmtn")
statustype <- RODBC::sqlFetch(ch0, "tbl_Status_Type")
welltype <- RODBC::sqlFetch(ch0, "Well_Type")

RODBC::odbcClose(ch0) 
```



```{r}

welldata <- welldata0 %>% 
  clean_names() %>% 
  rename(api = api_well_no) %>% 
  rename(status = wl_status) %>% 
  rename(twp = wh_twpn) %>% 
  rename(rng = wh_rng_n) %>% 
  rename(lat = wh_lat) %>% 
  rename(lon = wh_long) %>% 
  rename(sec = wh_sec)

wells <- welldata %>% 
  select(api, county, co_name, well_nm, well_typ, status, twp, rng, sec, dt_spud, dt_prod, dt_pa, lease_name, field, lat, lon) %>% 
  # Add grepl grep-logical 
  filter(grepl('CHEYENNE', as.factor(county), ignore.case = TRUE)) %>% 
  filter(well_typ != "SWD") %>% 
  unite(trs, c(twp, rng, sec), sep = "-") %>% 
   mutate(year = year(dt_spud) )

# write_csv(wells, "./data/tallgrass_wells.csv")
    

 
```
 

```{r}
tops <- tops %>% 
  clean_names()

tops2 <- tops %>% 
  rename(api=api_well_no) %>% 
  select(api, county, co_name, well_typ, wl_status, fmtn_cd, producing, dt_spud, dt_pa, well_name, field)

```

```{r eval=TRUE}
minyear <-  min(wells$year, na.rm=TRUE)

```

The following is a summary of wells drilled in Cheyenne County, Nebraska from the year `r minyear` to February 17, 2022.  The data is provided by Nebraska Oil & Gas Commission. ^[Nebraska Oil and Gas Commission, http://www.nogcc.ne.gov/Publications/NebraskaWellData.zip (last visited April 18, 2022)]


<br>  
<br> 

**Table 1.** Wells Drilled in Cheyenne County, Nebraska. 
```{r}
library(gt)

wells %>% 
  select(year, well_typ) %>% 
  group_by(well_typ) %>% 
  summarise(count = n()) %>%
  mutate(freq = round(count/sum(count),2)) %>% 
  arrange(desc(count)) %>% 
  gt() %>% 
  tab_header(
    title = "Count of Wells Drilled by type",
    subtitle = "Cheyenne County, NE"
  ) %>% 
  grand_summary_rows(
    columns = c(count, freq),
    fns = list(
      sum = ~sum(.)),
    formatter = fmt_number,
    decimals = 0,
    use_seps = FALSE
  ) %>% 
  tab_source_note(
    source_note = "Source: Nebraska Oil & Gas Commission, Feb. 17, 2022.")

``` 

<br>  
<br>  

```{r }

library(lubridate)
wells1 <- wells %>% 
 # mutate(year = as.factor(year(dt_spud) )) 
   mutate(year = year(dt_spud) )

# shag$year <- as.numeric(shag$year)  # transform year from character into numeric variable

# Making a histogram to assess data distribution
(wells.hist <- ggplot(wells1, aes(year, fill=well_typ)) + geom_histogram() + #theme.clean())
   theme_minimal()) +
  ggtitle("Histogram of Well Counts Drilled by Year\n Cheyenne Co. NE")
```

\newpage  
<br> 



<br> 
<br> 
<br> 



```{r fig.cap="Figure 2. Geom Density by well type - Areas add up to 1.0"}
# cyls <- as.factor(mtcars$cyl)
# https://r-charts.com/distribution/histogram-group-ggplot2/  

# Description of Geomdensity plots:  
# https://www.youtube.com/watch?v=FzfE8tfbpvQ 

ggplot(wells1, aes(x = year, fill = well_typ)) +
  geom_density(alpha = 0.7) +
  theme_grey() +  # Default
  theme(legend.position = "top") +
  ggtitle("Kernal Density Estimation Plot of Wells Drilled by Year\n Cheyenne Co. NE")
         # subtitle = "Kernal Density Plot") 
#  xlab("Dose (mg)") + ylab("Teeth length")
```
Sheather, S. J. and Jones, M. C. (1991). A reliable data-based bandwidth selection method for kernel density estimation. Journal of the Royal Statistical Society series B, 53, 683–690. doi: 10.1111/j.2517-6161.1991.tb01857.x. https://www.jstor.org/stable/2345597.

