---
title: "05 ToothGrowth"
author: "Robert C Cline Sr"
date: '2022-04-16'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE) 
library(kableExtra)
library(tidyverse)
library(here)
# library(readr) 
#
# sjplots --- 
library(sjPlot)
library(sjmisc)
library(sjlabelled)  

# tab_model(weevil.m)
```


```{r, include=FALSE} 

source("../scripts/theme.clean.R", local = knitr::knit_global())
# or sys.source("your-script.R", envir = knitr::knit_global())
```

<br>  
<br> 


#### **ToothGrowth** is a built-in dataset describing tooth growth in *Cavia porcellus* (South American guinea pigs) under different vitamin C treatments.  `data()`  

1. Are higher doses of vitamin C beneficial for tooth growth?  
  - see Figure 1.  
2. Does the method of administration (orange juice, OJ, or ascorbic acid, VC) influence the effect of the dose?  
  - See Figures 2 and 3. 
3. What would be the predicted tooth length of a guinea pig given 1 mg of vitamin C as ascorbic acid?  

* To identify datasets which are loaded type:  `data()`  

data(ToothGrowth)

**Table 1**. Summary of ToothGrowth dataset.  
```{r}
summary(ToothGrowth)  %>% 
  # library(kableExtra)
 
  kbl(caption = "OJ = Orange Juice; VC = ascorbic acid") %>% 
  kable_classic(full_width = F, html_font = "Cambria", position = "left")  
  
```

<br> 

**Table 2**. Sample OBS of ToothGrowth dataset.  
```{r} 

set.seed(123)
sample_n(ToothGrowth, 6)  %>% 
  # library(kableExtra)
 
  kbl(caption = "OJ = Orange Juice; VC = ascorbic acid") %>% 
  kable_classic(full_width = F, html_font = "Cambria", position = "left")  
  
``` 

```{r fig.cap='Figure 1. Tooth Growth as a function of Vitamin C dose'}
(tooth.p <- ToothGrowth %>% 
   filter(supp == "OJ") %>% 
   ggplot(aes(x = dose, y = len)) +
       ggtitle('') +
    geom_point(colour = "#483D8B") +
    geom_smooth(method = glm, colour = "#483D8B", fill = "#483D8B", alpha = 0.6) +
   # scale_x_continuous(breaks = c(1975, 1980, 1985, 1990, 1995, 2000, 2005)) +
    theme.clean() +
    labs(x = " ", y = "Length of Tooth Growth")) 
```



```{r eval=FALSE}
tooth.p + tooth.p2
```

<br> 

```{r fig.cap='Figure 2. Tooth Length and Supplement Type'}
 

library(ggplot2)

(teeth.p <- ggplot(ToothGrowth, aes(len, supp)) + 
    ggtitle('') +
    geom_boxplot(fill = "#CD3333", alpha = 0.8, colour = "#8B2323") +
    theme.clean() +  
    theme(axis.text.x = element_text(size = 12, angle = 0)) +
  labs(x = "Tooth Length", y = "Supplement")) +
  facet_grid(.~supp)


``` 

<br> 

```{r fig.cap='Figure 3. Tooth Growth Dependence on Dose and Supplement'}
ToothGrowth$dose=as.factor(ToothGrowth$dose)
ggplot(ToothGrowth, aes(x=dose,y=len))+geom_boxplot(aes(fill = dose))+
ggtitle('')+ facet_grid(.~supp)+
theme(axis.title.y = element_text(colour="gray20",size=12,angle=90,hjust=.5,vjust=1),
   axis.title.x = element_text(colour="gray20"),
plot.title = element_text(vjust=1.5,size = 12,colour="black"),
axis.text.x = element_text(colour="red",size=10,angle=45,hjust=.5, vjust=.5)) +
  labs(y = "Length (mm)") +
  labs(x = "dose (mg)")
```

What would be the predicted tooth length of a guinea pig given 1 mg of vitamin C as ascorbic acid?  
```{r results='asis'}
# comment=NA # to remove all hashes
teeth.m <- lm(len ~ dose + supp, data = ToothGrowth)
# summary(apples.m) 


```

<br> 

**Table 3.** Regression Summary for tooth growth in guinea pigs with Vitamin C supplementation.
```{r}
library(sjPlot)
library(sjmisc)
library(sjlabelled)  

tab_model(teeth.m)
```

<br> 

```{r}
d1 <- data.frame(dose = as.factor(1), supp="OJ")
predict(teeth.m, d1, interval = "prediction")
```


Predicted from the model *teeth.m*, a dose of 1 mg of vitamine C as ascorbic acid is expected 
result in a tooth length of `r predict(teeth.m, d1, inteval = "prediction")` mm.^[**Question**:  Why does my calculation differ from the prediction calcuated from Table 3?]  

to be 21.59 mm; 95% CI = 19.17  to 24.01.

12.46 + 9.13 = 21.59 mm
Min 12.46 + 6.71 = 19.17
Max 12.46 + 11.55 = 24.01  


